---
title: "Data Export"
subtitle: A dataset file exported from the pilot database
---

```{css, echo=FALSE}
th {
  white-space: nowrap;
}
```

The data file was created for the initial analysis using the 4Demon data.

## Data Availability

A tab-delimited file containing all major data from the database is available for download at the following location.

 * [pilot_4demon.tsv.gz](https://github.com/seafood-hazards/4demon-pilot/releases/download/v0.1.0/pilot_4demon.tsv.gz)

## Data File Summary

The file `pilot_4demon.tsv.gz` is a denormalized (flat) export of the relational database, designed to facilitate immediate analysis without requiring complex SQL joins.

## Column Definition
```{r setup, include=FALSE}
library(tibble)
```

This page describes all columns in the exported sediment data file `pilot_4demon.tsv.gz`. The file is produced by joining six database tables and selecting the columns listed below.

```{r}
tribble(
  ~`Column Name`,        ~`Source Table`, ~`Type`,    ~`Description`,

  # --- Measurement identifier ---
  "survey_seq_no",       "sediment",      "INTEGER",  "Original sequential survey value identifier — unique per row.",

  # --- Project ---
  "project_id",          "project",       "INTEGER",  "Surrogate key linking to the project table.",
  "project",             "project",       "TEXT",     "Project name (e.g. BAGGER, MONIT_SED, MMP_NS_RBINS-MUMM, PMPZ-DBII_PMPZ_COAST, UA_VAN_GRIEKEN).",
  "campaign_code",       "project",       "TEXT",     "Campaign code combining project prefix and year (e.g. 99_1979). Groups all samples from the same annual survey.",
  "year",                "project",       "INTEGER",  "Year of sampling (1971–2014).",
  "month",               "project",       "INTEGER",  "Month of sampling (1–12).",

  # --- Station ---
  "station_id",          "station",       "INTEGER",  "Surrogate key linking to the station table.",
  "station_code",        "station",       "TEXT",     "Station identifier code (162 unique stations).",
  "latitude",            "station",       "REAL",     "Corrected latitude of the sampling station (decimal degrees, WGS84).",
  "longitude",           "station",       "REAL",     "Corrected longitude of the sampling station (decimal degrees, WGS84).",
  "dist_to_coast",       "station",       "REAL",     "Distance from station to nearest coastline (km).",
  "est_country",         "station",       "TEXT",     "Estimated country derived from station coordinates.",
  "country_code",        "station",       "TEXT",     "ISO country code (e.g. BE for Belgium).",
  "municipality",        "station",       "TEXT",     "Municipality name nearest to the station.",
  "sea_name",            "station",       "TEXT",     "Name of the sea or marine region.",

  # --- Sample ---
  "sample_id",           "sample",        "INTEGER",  "Surrogate key linking to the sample table.",
  "gear_code",           "sample",        "TEXT",     "Sampling gear code: VV = Van Veen grab (bulk surface sediment), BC = box corer (intact core).",
  "depth_range",         "sample",        "TEXT",     "Sediment depth range sampled. Always 0–20 cm in this dataset.",

  # --- Method ---
  "method_id",           "method",        "INTEGER",  "Surrogate key linking to the method table.",
  "method_code",         "method",        "TEXT",     "Harmonised analytical method code combining project, technique, and fraction (e.g. Monit3_OES/MS_63 = monitoring period 3, ICP-OES/MS, <63 µm).",
  "matrix_code",         "method",        "TEXT",     "Sediment matrix code: FS = fine sediment (<63 µm fraction), US = unsieved / bulk sediment.",
  "fraction_range_um",   "method",        "TEXT",     "Grain size fraction analysed in micrometres (e.g. 0-63 = fine fraction <63 µm, 0-2000 = bulk sediment <2 mm).",

  # --- Sample replicate ---
  "replicate_number",    "sample",        "INTEGER",  "Replicate index for repeated measurements on the same sample (1, 2, or 3).",

  # --- Parameter ---
  "parameter",           "parameter",     "TEXT",     "Parameter code: metals (AL, AS, CD, CR, CU, FE, HG, NI, PB, ZN) and PCB congeners (CB28, CB52, CB101, CB118, CB138, CB153, CB180).",
  "parameter_type",      "parameter",     "TEXT",     "Parameter category: metal or PCB.",
  "parameter_name",      "parameter",     "TEXT",     "Full parameter name (e.g. mercury, zinc, PCB congener 153).",

  # --- Measurement ---
  "value",               "sediment",      "REAL",     "Original measured value in µg/g dry weight.",
  "corrected_value",     "sediment",      "REAL",     "Corrected value after quality control (e.g. below-detection substitutions). Recommended for analysis.",
  "unit",                "sediment",      "TEXT",     "Unit of measurement (µg/g dry weight for all rows).",

  # --- Quality flags ---
  "value_flag",          "sediment",      "INTEGER",  "Data quality flag: 0 = valid, 1 = suspect, 2 = below detection limit, 3 = invalid.",
  "det_limit_flag",      "sediment",      "INTEGER",  "Detection limit flag: 0 = above detection limit, 1 = at or below detection limit.",
  "range_check_flag",    "sediment",      "INTEGER",  "Range check flag: 0 = within expected range, 1 = outside expected range, NA = not checked.",
  "outlier_extreme_flag","sediment",      "INTEGER",  "Extreme outlier flag per parameter: 0 = normal, 1 = moderate outlier, 2 = extreme outlier.",
  "outlier_stdev_flag",  "sediment",      "INTEGER",  "Standard deviation outlier flag: 0 = within normal range, 1 = outlier based on standard deviation threshold."
)
```
