---
title: "EFSA Submission v2"
subtitle: "Data export mapping for the FHF marine sediment data model"
---

```{css, echo=FALSE}
th {
  white-space: nowrap;
}
```

The page presents the mapping between the [pilot database](db-schema.qmd) fields and the [EFSA Format v2](efsa-format-v2.qmd) columns.

::: {.callout-note}
The location columns this mapping draws on (`dist_to_coast`, the nearest country and
its code, `municipality`, `sea_name`) are **reserved but empty** in the pilot
database. The mapping below is the schema mapping, not a statement about coverage:
the values are derived at the clean stage, which recomputes and overwrites them. See
[DB Schema](./db-schema.qmd).
:::

::: {.callout-note}
### Note on Field Coverage
Fields marked **Missing** are not captured by the pilot database and must be supplied manually (or left blank) at submission time. **Fixed Text** fields are constant across the whole 4Demon pilot dataset (e.g. all values are dry weight).
:::

```{r}
library(dplyr)
library(tibble)

efsa_submission_v2 <- tribble(
  ~Column,              ~Optionality,    ~`Source Table`,   ~`Source Columns`,     ~`Fixed Text`,  ~Missing, ~Notes,
  "sampleId",           "Recommended",   "sample",          "sample_ref_code",     NA,             NA,       "Composite code already combining station, gear, and timestamp",
  "samplingStrategy",   "Recommended",   NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "pristine",           "Mandatory",     NA,                NA,                    NA,             "Yes",    "Pollution status of the sampling area is not recorded in DB",
  "collDate",           "Mandatory",     "sample",          "start_date",          NA,             NA,       NA,
  "placeColl",          "Mandatory",     "station",         "sea_name",            NA,             NA,       NA,
  "lat",                "Recommended",   "station",         "latitude",            NA,             NA,       NA,
  "lon",                "Recommended",   "station",         "longitude",           NA,             NA,       NA,
  "sedDepth",           "Mandatory",     "sample",          "depth_range",         NA,             NA,       "Always \"0-20 cm\" in this dataset — spans the 0-5cm/5-40cm boundary and needs manual bucketing",
  "traceEl",            "Mandatory",     "parameter",       "parameter_name",      NA,             NA,       "Pilot DB covers 10 metals + 7 PCB congeners; FHF v2 targets metals only",
  "value",              "Mandatory",     "sediment",        "value",               NA,             NA,       NA,
  "sd",                 "Optional",      NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "unit",               "Mandatory",     "sediment",        "unit",                NA,             NA,       NA,
  "resultType",         "Recommended",   NA,                NA,                    "dry weight",   NA,       "All 4Demon values are dry weight",
  "moisture",           "Recommended",   NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "convertedValue",     "Mandatory",     "sediment",        "corrected_value",     NA,             NA,       "µg/g dry weight is numerically equivalent to mg/kg dry weight; no conversion needed",
  "extractionClass",    "Mandatory",     NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "methAn",             "Mandatory",     "method",          "method_code",         NA,             NA,       NA,
  "validatedMethod",    "Recommended",   NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "LOD",                "Mandatory",     NA,                NA,                    NA,             "Yes",    "sediment.det_limit_flag records only a 0/1 flag, not a numeric detection limit",
  "LODUnit",            "Mandatory",     NA,                NA,                    NA,             "Yes",    "Not recorded in DB (see LOD)",
  "LOQ",                "Recommended",   NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "LOQUnit",            "Recommended",   NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "accLab",             "Mandatory",     NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "phSed",              "Optional",      NA,                NA,                    NA,             "Yes",    "Not recorded in DB",
  "sieve",              "Mandatory",     "method",          "fraction_range_um",   NA,             NA,       "Derived: 0-63 = <63µm sieved fraction, 0-2000 = unsieved/bulk",
  "bulkAnalysis",       "Mandatory",     "method",          "matrix_code",         NA,             NA,       "Derived: matrix_code = US indicates bulk/unsieved analysis",
  "ocSed",              "Mandatory",     NA,                NA,                    NA,             "Yes",    "Not recorded in DB; the pilot dataset covers only metals and PCB congeners",
  "notes",              "Recommended",   NA,                NA,                    NA,             "Yes",    "Open field, left blank"
)

efsa_submission_v2 <- efsa_submission_v2 %>%
  mutate(across(where(is.character), ~coalesce(.x, "")))
```

## Data Mapping for Submission

```{r}
efsa_submission_v2 %>%
  filter(Missing != "Yes") %>%
  select(-Missing)
```

## Full Data Mapping

```{r}
efsa_submission_v2
```

## Data Export Method

As the submission tool is an Excel file, the data must be mapped column by column rather than through a standard row-based export. This approach ensures that the structure of the submission file is not disrupted.
