---
title: "EFSA Submission v1"
subtitle: "Data export mapping for the EFSA reporting tool"
---

```{css, echo=FALSE}
th {
  white-space: nowrap;
}
```

The page presents the mapping between the [pilot database](db-schema.qmd) fields and the [EFSA Format v1](efsa-format.qmd) columns.

::: {.callout-note}
The location columns this mapping draws on (`dist_to_coast`, the nearest country and
its code, `municipality`, `sea_name`) are **reserved but empty** in the pilot
database. The mapping below is the schema mapping, not a statement about coverage:
the values are derived at the clean stage, which recomputes and overwrites them. See
[DB Schema](./db-schema.qmd).
:::

::: {.callout-note}
### Note on Field Coverage
Fields marked **Missing** are not captured by the pilot database and must be supplied manually (or left blank) at submission time. **Fixed Text** fields are constant across the whole 4Demon pilot dataset (e.g. all values are dry weight, all samples are marine sediment).
:::

```{r}
library(dplyr)
library(tibble)

efsa_submission <- tribble(
  ~Column,           ~Lookup, ~Optionality,           ~Table,      ~`Column Names`,               ~`Fixed Text`,                                   ~Missing,
  "sampleId",        NA,      "Mandatory",            "sample",    "sample_ref_code",              NA,                                              NA,
  "recId",           "Yes",   "Mandatory",            NA,          NA,                             NA,                                              NA,
  "envComp",         NA,      "Mandatory",            NA,          NA,                             "marine sediment",                               NA,
  "pristineLoc",     NA,      "Mandatory",            NA,          NA,                             NA,                                              "Yes",
  "sampleLocGC",     NA,      "Mandatory",            "station",   "latitude + longitude",         NA,                                              NA,
  "sampleLocCM",     NA,      "Mandatory",            "station",   "est_country + municipality",   NA,                                              NA,
  "sampleDate",      NA,      "Mandatory",            "sample",    "start_date",                   NA,                                              NA,
  "traceElText",     NA,      "Mandatory",            "parameter", "parameter_name",               NA,                                              NA,
  "traceElCode",     "Yes",   "Mandatory",            NA,          NA,                             NA,                                              NA,
  "specText",        NA,      "Mandatory",            "parameter", 'parameter_name + " (Total)"',  NA,                                              NA,
  "specCode",        "Yes",   "Mandatory",            NA,          NA,                             NA,                                              NA,
  "conc",            NA,      "Mandatory",            "sediment",  "corrected_value",              NA,                                              NA,
  "unitText",        NA,      "Mandatory",            "sediment",  "unit",                         NA,                                              NA,
  "unitCode",        "Yes",   "Mandatory",            NA,          NA,                             NA,                                              NA,
  "weightText",      NA,      "Dependent Mandatory",  NA,          NA,                             "dw",                                            NA,
  "weightCode",      "Yes",   "Dependent Mandatory",  NA,          NA,                             NA,                                              NA,
  "methAnText",      NA,      "Mandatory",            "method",    "method_code",                  NA,                                              NA,
  "methAnCode",      "Yes",   "Mandatory",            NA,          NA,                             NA,                                              NA,
  "LOD",             NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "LOQ",             NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "accLab",          NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "phSed",           NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "TextureSedClay",  NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "TextureSedSilt",  NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "TextureSedSand",  NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "ocSed",           NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "phWater",         NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "hardWater",       NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "DOC",             NA,      "Optional",             NA,          NA,                             NA,                                              "Yes",
  "publicData",      NA,      "Optional",             NA,          NA,                             "Yes",                                           NA,
  "refPublication",  NA,      "Optional",             NA,          NA,                             "https://www.vliz.be/projects/4demon/index.htm", NA,
  "comments",        NA,      "Optional",             NA,          NA,                             NA,                                              "Yes"
)

efsa_submission_full <- efsa_submission %>%
  mutate(
    across(
      where(is.character),
      ~coalesce(.x, "")
    )
  ) %>%
  rename(c(`Source Table` = "Table",
           `Source Columns` = "Column Names"))

efsa_submission <- efsa_submission_full %>%
  filter(Lookup != "Yes") %>%
  dplyr::select(-c(Lookup, Missing))

```

## Data Mapping for Submission
```{r}
efsa_submission
```

## Full Data Mapping
```{r}
efsa_submission_full
```

## Data Export Method
As the EFSA submission tool is an Excel file containing embedded macros, the data must be mapped column by column rather than through a standard row‑based export. This approach ensures that the structure of the submission file is not disrupted.

Note that the EFSA v1 catalogue's trace element list (Co, Cu, I, Mn, Mo, Se, Zn) only partially overlaps with the pilot database's `parameter` table (10 metals: AL, AS, CD, CR, CU, FE, HG, NI, PB, ZN, plus 7 PCB congeners). Only copper (`CU`) and zinc (`ZN`) can currently be submitted under this format.
